---
title: "HW 6"
author: "Nandini M"
date: "2023-12-02"
output: github_document
  #html_document:
    #toc: true
    #toc_float: true
    #code_folding: hide
---

Load in libraries and set up display settings
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)
library(modelr)
library(mgcv)
library(readxl)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
set.seed(123)
```

## Problem 2 

Load in Data:

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data: r^2, log(β^1∗β^2)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 

```{r}
boot_strap_weather = weather_df |> 
  modelr::bootstrap(n = 5000, id = "strap_number") |> 
  mutate(
    models = map(.x = strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    corr = map(models, broom::glance)) |> 
  select(-strap, -models) |> 
  unnest(results) |> 
  filter(term %in% c("tmin","prcp")) |> 
  group_by(strap_number) |> 
  mutate(log_b1x2 = log(sum(estimate))) |> 
  select(log_b1x2, corr) |> 
  unnest(corr) |> 
  janitor::clean_names() |> 
  select(strap_number, log_b1x2, r_squared)

boot_strap_weather
```

Plot the distribution of your estimates, and describe these in words. 

```{r}
corr_dist = boot_strap_weather |> 
  ggplot(aes(x = r_squared)) + geom_density()

print(corr_dist + ggtitle("Density Plot of r^2 Sample Estimates"))

log_dist = boot_strap_weather |> 
  ggplot(aes(x = log_b1x2)) + geom_density()

print(log_dist + ggtitle("Density Plot of log(β^1xβ^2 Sample Estimates"))
```

Across the 5000 bootstrap samples, the value of the sample r-squared estimate approximately ranges between 0.915 to 0.93. There is a high proportion of samples with a correlation coefficient close to 1.

Across the 5000 bootstrap samples, the values of the sample log(β^1∗β^2) estimate approximately ranges between 0.012 to 0.025.

Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^1∗β^2).

```{r}

```

